---
title: "EDA contd"
author: "fatima"
date: "July 2020"
output: html_document
---


```{r}
library(gridExtra)
library(ggplot2)
library(dplyr)
library(corrplot)


data <- read.csv("C:/Users/fatim/Downloads/capstone initial results/df.csv",stringsAsFactors = TRUE)
str(data)
```

## dropping first column i.e index column


```{r}
data<-data[-1]
data<-transform(data,arrival.year = factor(arrival.year),repeated.guest = factor(repeated.guest),    is.cancelled = factor(is.cancelled))
data$last.reservation.status.date <- as.Date(data$last.reservation.status.date,format=)
data$country <- as.character(data$country)
data$agent <- as.character(data$agent)
```



```{r}

## ordering arrival.month variable

data$arrival.month<- factor(data$arrival.month,levels=month.name)

```


```{r}
## creating column total nights and total guests 

data<-mutate(data,total.nights=total.weekend.nights+total.week.nights)
data<-mutate(data,total.guests=adults+children+babies)
```


## filtering each year data
```{r}
data.2015 <-filter(data,arrival.year==2015)
data.2016<-filter(data,arrival.year==2016)
data.2017<-filter(data,arrival.year==2017)

```



```{r}

## to see which month has highest nights booking in 2015,2016,2017
## for three years busiest months are July and august but cancellations have increased in year 2017


p1<-data.2015 %>% group_by(arrival.month,is.cancelled) %>% summarise(total_nights = sum(total.nights)) %>% 
ggplot(aes(x = arrival.month, y = total_nights, fill = is.cancelled)) +
geom_bar(stat = "identity",position='dodge',width=0.5) +labs( x = "Month", y = "total nights",title = ("Total nights booked in 2015")  )


## for 2016
p2<-data.2016 %>% group_by(arrival.month,is.cancelled) %>% summarise(total_nights = sum(total.nights)) %>% 
ggplot(aes(x = arrival.month, y = total_nights, fill = is.cancelled)) +
  geom_bar(stat = "identity",position='dodge',width=0.5) +theme(axis.text.x = element_text(angle = 45,hjust = 1,vjust = 0.5))+
labs(x = "Month",y = "total nights",title = ("Total nights booked in 2016")  )

## for 2017
p3<-data.2017 %>% group_by(arrival.month,is.cancelled) %>% summarise(total_nights = sum(total.nights)) %>% 
ggplot(aes(x = arrival.month, y = total_nights, fill = is.cancelled)) +
  geom_bar(stat = "identity",position='dodge',width=0.5) +
labs(x = "Month",y = "total nights",title = ("Total nights booked in 2017")  )

grid.arrange(p1, p2,p3 ,nrow=3,heights=c(0.6,0.75,0.6))

```


```{r}
## guest with family or adult only
## less bookings are with families- only 3869

only.family.kids<-filter(data,total.guests > adults)
with.no.kids<-filter(data,total.guests==adults)
nrow(only.family.kids)
nrow(with.no.kids)
```


```{r}
## guests with family mainly visit during july and august and they tend to cancel more as compared to adults only guests


b<-only.family.kids %>% group_by(arrival.month,is.cancelled) %>% summarise(total_nights = sum(total.nights))
plot1<-ggplot(b,aes(x = arrival.month, y = total_nights, fill = is.cancelled)) +
               geom_bar(stat = "identity",position='dodge',width=0.5) +
                  theme(axis.text.x = element_text(angle = 45,hjust = 1,vjust = 0.5))+
                   labs(x = "Month",y = "total nights",title = ("Total nights booked by guest with family ")  )

d<-with.no.kids %>% group_by(arrival.month,is.cancelled) %>% summarise(total_nights = sum(total.nights))
plot2<-ggplot(d,aes(x = arrival.month, y = total_nights, fill = is.cancelled)) +
    geom_bar(stat = "identity",position='dodge',width=0.5) +
      theme(axis.text.x = element_text(angle = 45,hjust = 1,vjust = 0.5))+
        labs(x = "Month",y = "total nights",title = ("Total nights booked by guest with no kids ") ) 

grid.arrange(plot1, plot2, nrow=2,heights=c(0.6,0.6))
```



```{r}
## visualising above
## guests with families cancel more if total guest are 4(almost 50% chance)
## guests with adults only cancel ratio is approx 1/5
## guests with family cancel ratio is approx 1/3


plot1<-ggplot(only.family.kids,aes(x= factor(total.guests)))+geom_bar(aes(fill=is.cancelled))
plot2<-ggplot(with.no.kids,aes(x= factor(adults)))+geom_bar(aes(fill=is.cancelled))
grid.arrange(plot1, plot2, ncol=2)

```


```{r}

## looking whether change in room type result in cancellation
## 7006 observations have different assigned room than what they reserved
##  change in room type did not have any  effect on cancellation


roomtype.diff<-filter(data,as.character(assigned.room.type) != as.character(reserved.room.type))
roomtype.same<-filter(data,as.character(assigned.room.type) == as.character(reserved.room.type))
prop.table(table(roomtype.diff$is.cancelled))
prop.table(table(roomtype.same$is.cancelled))
```

```{r}
## creating colum to find percentage of people coming from different countries
## PRT has approx. 43% of total booking


country.data<-data %>% group_by(country) %>% summarise(country.count = n())
country.data<-mutate(country.data,country.percent=(country.count/sum(country.count))*100)
country.greater.2percent<-filter(country.data,country.percent>2)


```


## top three countries are PRT,GBR,ESP
```{r}
ggplot(country.greater.2percent, aes(x = "", y =country.percent, fill = country)) +
  geom_col(width = 1,color = "white") +
  coord_polar("y", start = 0)+
   ggtitle("percentage of Country greater than 2%")+
  theme_void()

```
## replacing null with PRT
## 1/3 of PRT booking was cancelled
```{r}
data$country<-replace(data$country,data$country=='NULL','PRT')
country.data.new<-data %>% group_by(country,is.cancelled) %>% summarise(country.count = n())
country.data.top3<-filter(country.data.new,country=="PRT" | country=="ESP"|country=="GBR")
qplot(country,country.count,data=country.data.top3,shape=is.cancelled,color=is.cancelled,size=country.count)

```

## grouping country variable
```{r}
group_country<-function(countryname){
  
  if(countryname =='PRT'){
    return('High Frequency')
  }else if(countryname == "ESP" | countryname=="GBR"){
    return('Medium Frequency')
  }else{
    return('Low Frequency')
  }
}

```
## grouping country and creating new column of grouped data
```{r}
data$grp.country<-sapply(data$country,group_country)
data$grp.country<-factor(data$grp.country)
levels(data$grp.country)

```

```{r}
## changing levels so that they are in order

data$grp.country<-factor(data$grp.country,levels=c("High Frequency","Medium Frequency","Low Frequency"))
levels(data$grp.country)
table(data$grp.country)
```

```{r}
## extracting only year from last.reservation.status.date variable
last.res.date.year<-substring(as.character(data$last.reservation.status.date),1,4)## extracting only year
last.res.date.year<-factor(last.res.date.year)
data$last.reservation.status.year<-last.res.date.year

```




##finding variance of numeric variables
```{r}
only.num<-data %>% Filter(f=is.numeric) %>% colnames()
v1<-round(sapply(data[only.num],var),2)%>%sort()
v<- as.data.frame(v1)
v
```

```{r}
## creating a vector for numeric variables  with var>1.5)
feature.model.numeric<-c("total.week.nights","total.nights","days.waiting.list","arrival.day.of.month","arrival.week.of.year","adr","lead.time")

```


```{r}
## checking correlation for selected numeric variables,method used is spearman as data is not normal and has outliers

d1<-cor(data[feature.model.numeric],method="spearman")
corrplot(d1)

```

```{r}
## keeping only one from total.nights or total.week.nights as both are highly correlated
## removing total.week.nights as its variance is low compared to total.nights


feature.model.numeric<-feature.model.numeric[-1]
feature.model.numeric
```

## treatment of outliers of numeric variables by creating groups and then changing to factor
## for lead.time
```{r}
group_lead.time<-function(time){
  
  if(time<20){
    return('days elapsed <20')
  }else if(time>=20 & time<100){
    return('20>=days elapsed <100')
  }else{
    return('days elapsed >100')
  }
}


```

```{r}
data$grp.lead.time<-sapply(data$lead.time,group_lead.time)
data$grp.lead.time<-factor(data$grp.lead.time)
levels(data$grp.lead.time)

```

```{r}
data$grp.lead.time<-factor(data$grp.lead.time,
                           levels=c("days elapsed <20","20>=days elapsed <100","days elapsed >100"))
levels(data$grp.lead.time)

table(data$grp.lead.time)
```
## treating outliers of adr by creating groups
```{r}
group_adr<-function(amt){
  
  if(amt<45){
    return('very reasonable')
  }else if(amt>=45& amt<65){
    return('reasonable')
  }else if(amt>=65& amt<95){
    return('ok')
  }else if(amt>=95& amt<155){
    return('high')
  }else{
    return('pretty high')
  }
}
```




```{r}
## removing records where adr=0 and total nights =0 for more generalised model

data<-filter(data,adr!=0 &total.nights!=0)
```

```{r}
data$grp.adr<-sapply(data$adr,group_adr)
data$grp.adr<-factor(data$grp.adr)
levels(data$grp.adr)

```

```{r}
data$grp.adr<-factor(data$grp.adr,levels=c("very reasonable","reasonable","ok","high","pretty high"))
levels(data$grp.adr)
table(data$grp.adr)

```


## treating outliers of total.nights by creating groups

```{r}
group_nights<-function(num){
  
  if(num<3){
    return('less than 3')
  }else if(num>=3& num<6){
    return('3 or more but less than 6')
  }else{
    return('6 or more')
  }
}

data$grp.total.nights<-sapply(data$total.nights,group_nights)
data$grp.total.nights<-factor(data$grp.total.nights)
levels(data$grp.total.nights)

```

```{r}
data$grp.total.nights<-factor(data$grp.total.nights,levels=c("less than 3","3 or more but less than 6","6 or more"))
levels(data$grp.total.nights)
table(data$grp.total.nights)

```



```{r}

## replacing undefined  with SC in meal as both are same 
## combining group customer type  to contract type
## renaming level contract to other


data$meal<-replace(data$meal,data$meal=='Undefined','SC')
data$meal<-factor(data$meal,levels=c("BB","FB","HB","SC"))
levels(data$customer.type)<-c("Contract","Contract","Transient","Transient-Party")
levels(data$customer.type)[1]<-"Other"
table(data$customer.type)
```

```{r}
## combining no-show with canceled in last.reservation.status

levels(data$last.reservation.status)<-c("Canceled","Check-Out","Canceled")

table(data$last.reservation.status)
```

## Feature selection

```{r}
## removing numeric variables with var<1.5 
## Also removing lead.time,adr,total.nights( will be using the grouped one of these variables)


data.new<-data[-c(1,6,7,8,9,10,16,17,20,25,26,27,31,32)]
```



```{r}

## normalising numeric variables


fun<- function(x){
  x<- (x-min(x))/(max(x)-min(x))
  return(x)
}
new.numeric <-c("days.waiting.list","arrival.day.of.month","arrival.week.of.year")
nor.data<-sapply(data.new[,new.numeric],fun)
nor.data<-as.data.frame(nor.data)
```

```{r}
## selecting factor variables

only.fact<-data.new %>% Filter(f=is.factor) %>% colnames()
only.fact
```



```{r}
##removing country,agent and last.reservation.status.date

data.new1<-data.new[-c(6,13,17)]
```


 
```{r}

## changing factor to int to find correlation

data.final.corr<-data.new1
factor.final.names = data.final.corr %>% Filter(f=is.factor) %>% colnames()
data.final.corr.copy<-data.final.corr
data.final.corr.copy[,factor.final.names] = data.frame(sapply( data.final.corr.copy[,factor.final.names], as.integer))
str(data.final.corr.copy)

```
 
## finding correlation
 
```{r}
d2<-cor(data.final.corr.copy,method="spearman")

corrplot(d2,method = "color")
```


```{r}
## based on  corrplot and bar plot of factor variables removing following variables because of correlation or because of less information
##last.reservation.status.year,arrival.week.of.year,distribution.channel,assigned.room.type,
## repeated guest,deposit.type,days.waiting.list

```


```{r}
data.new1<-data.new1[-c(17,3,7,10,8,11,12)]
colnames(data.new1)
data.model1<-data.new1
```


```{r}
## also removing last.reservation.status as it is almost similar to target variable

data.model1<-data.model1[-8]

```


```{r}
## from reserved.room.type removing level B and L as they have very few entries and it might be problem for machine learning because it might give zero frequency when partioning dataset


xtabs(~reserved.room.type+is.cancelled,data.model1)
data.model1<-filter(data.model1,reserved.room.type!="B" & reserved.room.type!="L")
data.model1$reserved.room.type<-droplevels(data.model1$reserved.room.type)
xtabs(~reserved.room.type+is.cancelled,data.model1)
```

```{r}

#data.model.copy<-write.csv(data.model1,"C:/Users/fatim/Downloads/capstone initial results/data.for.model.csv")
```






