---
title: "machine learning"
author: "fatima"
date: "28/06/2020"
output: html_document
---



##install.packages("caret")for evaluating data
##install.packages("e1071")for confusion matrix
```{r}
library(caret)
library(e1071)
library(ROSE)
library(randomForest)
data.model <- read.csv("C:/Users/fatim/Downloads/capstone initial results/data.for.model.csv")
str(data.model)
```


```{r}
data.model<-data.model[-1]
data.model$arrival.year<-as.factor(data.model$arrival.year)
data.model$is.cancelled <-as.factor(data.model$is.cancelled )
```


## splitting dataset into training and testing
```{r}
set.seed(101)
row.no.train<-sample(1:nrow(data.model),nrow(data.model)*0.8)
train.hotel<-data.model[row.no.train,]
test.hotel<-data.model[-row.no.train,]
```


#removing target variable from test data

```{r}
test.hotel.new<-test.hotel[-8]
```

## random forest
```{r}
model.rf.tree<-randomForest(is.cancelled ~ .,data=train.hotel)
predict.rf<-predict(model.rf.tree,test.hotel.new)
table(predict.rf,test.hotel$is.cancelled)
```

```{r}
confusionMatrix(predict.rf,test.hotel$is.cancelled,positive="1")
```


## balancing dataset(only training set)

```{r}
table(train.hotel$is.cancelled)
prop.table(table(train.hotel$is.cancelled)) ## checking proprotion of levels of is.cancelled
```


## oversampling
```{r}
over.data<- ovun.sample(is.cancelled~.,data=train.hotel,method="over",N=40730)$data
table(over.data$is.cancelled)

```

## building model

```{r}
model.rf.over.balanced<-randomForest(is.cancelled ~ .,data=over.data)
confusionMatrix(predict(model.rf.over.balanced,test.hotel.new),test.hotel$is.cancelled,positive="1")
```
## undersampling
```{r}
under.data<-ovun.sample(is.cancelled~.,data=train.hotel,method="under",N=13304)$data
table(under.data$is.cancelled)
```
## model with undersampling
```{r}
model.rf.under.balanced<-randomForest(is.cancelled ~ .,data=under.data)
confusionMatrix(predict(model.rf.under.balanced,test.hotel.new),test.hotel$is.cancelled,positive="1")
```
## over and under both
```{r}
both.data<-ovun.sample(is.cancelled~.,data=train.hotel,method="both",seed=123,N=26608)$data
table(both.data$is.cancelled)
```

```{r}
model.rf.both.balanced<-randomForest(is.cancelled ~ .,data=both.data)
confusionMatrix(predict(model.rf.both.balanced,test.hotel.new),test.hotel$is.cancelled,positive="1")
```

## Random forest with imbalanced dataset-> accuracy is 82.44%
## After oversampling -> accuracy is 79.9%
